---
title: "Question 2"
output: github_document
---
```{r setup, echo=FALSE, message=FALSE, comment=NA}
library(mosaic)
brca <- read.csv("~/Desktop/data mining 作业/data/brca.csv")
```
Let's model recall versus risk factors + radiologist, the coefficients of each variables are shown as below. 
From this generalized model, we can see for radiologist 34, he has exp(-0.52) ≈ 0.6 to recall a patient, holding all other factors constant.
radiologist 89 has exp(0.46) ≈ 1.58 times to recall a patient, holding all other factors constant.
radiologist66 has exp(0.35) ≈ 1.42 times to recall a patient, holding all other factors constant.
So we can say radiologist 89 is more conservative while radiologist 34 is less conservative among radiologists, holding all other factors constant.
```{r, echo=FALSE, comment=NA}
model_recall = glm(recall ~ . - cancer, data=brca, family=binomial)
coef(model_recall)
```
Then we can consider which radiologist is more accurate by seeing the confusion matrix.
For FDR, for radiologist 89, it is 33/38, while for radiologist 34, it is 13/17, which the radiologist 89 has the higher one.
However, for FDR, it is lower for radiologist 89, as 2/7 is lower than 3/7 for radiologist 34.
So considering the confusion matrix, it is not clear to say for the accuracy.
```{r, echo=FALSE, comment=NA}
xtabs(~cancer + recall, data=subset(brca, radiologist == 'radiologist34'))
xtabs(~cancer + recall, data=subset(brca, radiologist == 'radiologist89'))
```
For part 2 in this question.
Firstly, for Model A, we regress the cancer outcome on the recall decision. The coefficients and the significance level of the model is shown as below.
We can say that as recall increase by 2 unit, the cancer will increase by exp(2.26) = 9.58.
```{r, echo=FALSE, comment=NA}
model_cancer1 = glm(cancer ~ recall, data=brca, family=binomial)
coef(model_cancer1)
summary (model_cancer1)
```
And then we regress the Model B for the cancer on the recall decision and family history.The coefficients and the significance level of the model is shown as below.
```{r, echo=FALSE, comment=NA}
model_cancer2 = glm(cancer ~ recall + history, data=brca, family=binomial)
coef(model_cancer2)
summary (model_cancer2)
```
If the radiolist were appropriately accounting for a patients' family history for the cancer, I think the regression result of Model B would be better than Model A.
Because for model B, it takes the ommited variable in Model A into consideration.
However, by seeing the p-value, AIC index and the R-squared, Model B is not better than Model A, because AIC for Model B is higher and p-value is not significant.
From this result, we can say for radiologist, the family history is not a main reason for this cancer to justify the existence of cancer and recall.
After that, we should also consider other reasons for radiologists to recall patients.
To be specific, we can model the cancer result on the risk factors
```{r, echo=FALSE, comment=NA}
model_cancer = glm(cancer ~ ., data=brca, family=binomial)
coef(model_cancer)
```
From the regression result, we can see that patients and older have exp(1.44) ≈ 4.2 the probability of having cancer, holding all else fixed.
So it is reasonable for the radiologists to pay more attention to the patients who are 70 and older.
Also, patients with tissue density classification 4 have exp(2) ≈ 7.4 times the probability of having cancer as patients with density 1, holding all else fixed.
For radiologists, in order to make decisions more precisely, we should also consider the error rates, shown as below.
For patiens who are 70 and older, the FDR, it is 70%. And for the patients with density 4, the FDR is 0.73.
```{r, echo=FALSE, comment=NA}
xtabs(~cancer + recall + age, brca) %>% prop.table(margin=c(2, 3))
xtabs(~cancer + recall + density, brca) %>% prop.table(margin=c(2, 3))
```