---
title: "Exercise1-1 XuechunWang"
output: md_document
---

Exercise 1-1
========================================================
- Course: Data Mining and Statistical Learning (ECO395M)
- Name: Xuechun Wang (xw5996)
- Date: Feburary 7th, 2020
- Data Source: ABIA.csv

What is the best time to fly from/arrive at Austin Airport?
========================================================

I'm goin to show a set of plots here trying to help traveller and those who always need to take a flight to figure out what is the best time to fly from/arrive at Austin to minimum delays.

Best hour to fly from Austin in a day
========================================================

In general, the flights that departed from Austin at night had longer delays, especially after 18:00. There is a significant trend of increasing delays as it got later in a day. The delay during the daytime(5:00-18:00) is not serious, and the overall delay is at a relatively low level, but the flight at noon time(11:00-14:00) had relatively longer delays than the flights in the morning(5:00-11:00) and afternoon(14:00-18:00). Flights departed from Austin during the morning had the shortest delay. So if you need to take a flight out of Austin, a morning flight will be the best choice.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(mosaic)
library(tidyverse)
library(readr)
ABIA <- read_csv("ABIA.csv")
DepAus = subset(ABIA, Origin== 'AUS')
ArvAus = subset(ABIA, Dest== 'AUS')

library(stringr)
library(data.table)
library(magrittr)

DepAus <- 
  DepAus %>% 
  mutate(
    Hour=DepTime %>% 
    str_sub(.,end=nchar(.)-2)%>% 
    as.numeric
  )

ggplot(data=DepAus) + 
  geom_boxplot(aes(group=Hour, x=Hour, y=DepDelay, )) + 
  theme_bw(base_size=18)+xlim(4,25)+ylim(0,300)
```

Best month to fly from Austin in a year
========================================================

The difference between months is not significant. The delays in September were minimal. The overall delays in September, October, and November were less, but the delays began to increased significantly in December. Due to the arrival of the holiday season, people's travel increased, and the number of aircraft taking off at the airport in one day increased, which made the delay time increase. In addition, the overall delay in winter is relatively long, which is also related to the extreme weather. For all other months, April had shorter delays in average. In conclusion, if you want to choose a time to travel out of Austin, Autumn(Sep-Nov) will be the best time.

```{r, echo=FALSE, warning=FALSE}
ggplot(data=DepAus) + 
  geom_boxplot(aes(group=Month, x=Month, y=DepDelay)) + 
  theme_bw(base_size=18)+ ylim(0,300) +scale_x_continuous(breaks = seq(1,12,1))
```

Scatter plot (faceting on the third variable: Month)
========================================================

In different month group, the relationship between Departure hour and Delay time didn't have significant differences. For most of the month, it would be better to take a morning flight or even a afternoon flight than the flight departed at noon or at night.

```{r, echo=FALSE, warning=FALSE}
ggplot(data = DepAus) + 
  geom_point(mapping = aes(x = Hour, y = DepDelay)) + 
  facet_wrap(~ Month, nrow = 6) + ylim(0,300)
```

Best hour to arrive at Austin in a day
========================================================

In general, the flights that arrived at Austin at night had longer arrival delays, especially after 20:00. There is a significant trend of increasing delays as it got later in a day. The delay during the daytime(5:00-19:00) is not serious, and the overall delay is at a relatively low level. Flights arrived at Austin during the morning had the shortest delay. So if you need to take a flight to Austin, a flight arrived at morning will be the best choice.

```{r, echo=FALSE, warning=FALSE}
ArvAus <- 
  ArvAus %>% 
  mutate(
    Hour=DepTime %>% 
    str_sub(.,end=nchar(.)-2)%>% 
    as.numeric
  )

ggplot(data=ArvAus) + 
  geom_boxplot(aes(group=Hour, x=Hour, y=ArrDelay )) + 
  theme_bw(base_size=18)+ylim(0,400)
```

Best month to arrive at Austin in a year
========================================================

The difference between months is not significant. The delays in October were minimal. The overall delays in September, October, and November were less, but the delays began to increased significantly in December. For all other months, April had shorter delays in average. And the overall Arrival Delay had very similar trend with the Departure Delay, which make me believe that there is a strong connection between the Departure Delay and Arrival Delay. And the Delay trend didn't just apply to Austin Airport but all other airports in the U.S. In conclusion, if you want to choose a time to travel to Austin, Autumn(Sep-Nov) will be the best time to minimize your delay when taking flights.

```{r, echo=FALSE, warning=FALSE}
ggplot(data=ArvAus) + 
  geom_boxplot(aes(group=Month, x=Month, y=ArrDelay)) + 
  theme_bw(base_size=18)+ ylim(0,300) +scale_x_continuous(breaks = seq(1,12,1))
```

Scatter plot (faceting on the third variable: Month)
========================================================

In different month group, the relationship between Arrival hour and Delay time didn't have significant differences. For most of the month, it would be better to take a flight arrived at Austin in morning to minimize delays.

```{r, echo=FALSE, warning=FALSE}
ggplot(data = ArvAus) + 
  geom_point(mapping = aes(x = Hour, y = ArrDelay)) + 
  facet_wrap(~ Month, nrow = 6) + ylim(0,300)
```

Summary
========================================================

In Conclusion, the best time to fly from Austin has very similar trends with the best time to arrive at Austin, which implies that there is a strong connection between Departure Delays and Arrival Delays. And it also implies that this trend applies to most of the U.S. airport (because the arrival flights flied to Austin from other cities). 
For those who traveled a lot by flights, the best time to fly from/arrive at Austin will be morning. And for those who are planning a trip to Austin or are trying to travel to somewhere else from Austin, Autumn will be the best time of a year to minimize flight delays.




